Methods for understanding the variable importance of local explanations of black-box models

### 2021-11-03 _paper.rmd
classifciation task.

Data visualization tours animate many linear projections over small changes to the basis. The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame. Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame. The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution. In the next sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.

@@

@@

@@

<!-- render time -->
The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with __shiny__ [@chang_shiny_2021]. The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievet_interactive_2020]. The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.

>>>>

Data visualization tours animate many linear projections over small changes to the basis. The manual tour creates a basis path by varying the contribution of a selected variable, fully into and out of a projection frame. Doing allows an analyst to test an individual variable's sensitivity to the structure identified in the frame. The default variable selected is the one with the largest discrepancy between the primary and comparison observation's attribution. In the following sections we elaborate on the takeaways we draw from applying this approach in classification and regression tasks respectively.

@@

What information do we glean from using this method on a classifcation task?  Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space. We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction. By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation. That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.

Another way of thinking about this would be: what would the predictions be if the explanation were different. This is similar to the idea of _ceteris paribus_ profiles [@biecek_ceterisparibus_2020]. _Ceteris paribus_ is Latin for “other things held constant” or “all else unchanged”. The profiles visualize what 'what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant. In contrast to _ceteris paribus_ profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.

@@

The regression case is not as discrete a feature. Instead the prediction or the residual becomes the comparison of accuracy.

@@

<!-- render time -->
The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with __shiny__ [@chang_shiny_2021]. The global view is created first with __ggplot2__ [@wickham_ggplot2_2016] which is then rendered to an interactive html widget with __plotly__ [@sievert_interactive_2020]. The tour view is created in __spinifex__ [@spyrison_spinifex_2020], which creates the manual tour basis array, and facilitates similar rendering to html widget.

### 2021-11-03 _paper.tex
classifciation task.

@@

The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with \textbf{shiny} (Chang et al. 2021). The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (\textbf{sievet\_interactive\_2020?}). The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.

@@

>>>>

What information do we glean from using this method on a classifcation task? Typically we select a misclassified observation in comparison with correctly classified point that is nearby in data space. We start by seeing the data projected through the linear attribution, the combination that best justifies that prediction. By default the manual tour varies the contribution of the variable with the largest difference between the primary and comparison observation. That is, we can test the sensitivity of each variable to structure identified by the local explanation, we are exploring the support of the explanation, evaluating the support or robustness of the prediction.

Another way of thinking about this would be: what would the predictions be if the explanation were different. This is similar to the idea of \emph{ceteris paribus} profiles (\textbf{biecek\_ceterisparibus\_2020?}). \emph{Ceteris paribus} is Latin for ``other things held constant'' or ``all else unchanged.'' The profiles visualize what `what-if' analysis, showing how an observation's prediction would change from a change in one explanatory variable given that other other variables are all held constant. In contrast to \emph{ceteris paribus} profiles, touring methods visualize all observations rather than one, and variables are not treated as independent, that is the projection has an orthonormal basis; when one variable is rotated out-of-frame, other variables are effectively rotated into-the-frame.

@@

The work remaining at runtime consists mostly in rendering the frames of the tour as specified by the selection of the parameters. The application is made with \textbf{shiny} (Chang et al. 2021). The global view is created first with \textbf{ggplot2} (Wickham 2016) which is then rendered to an interactive html widget with \textbf{plotly} (Sievert 2020). The tour view is created in \textbf{spinifex} (Spyrison and Cook 2020), which creates the manual tour basis array, and facilitates similar rendering to html widget.

@@

\leavevmode\hypertarget{ref-sievert_interactive_2020}{}%
Sievert, Carson. 2020. \emph{Interactive {Web}-{Based} {Data} {Visualization} with {R}, Plotly, and Shiny}. Chapman; Hall/CRC. \url{https://plotly-r.com}.
